### 2023-10-25 paper.tex
first commit

\documentclass{article}

\title{Dissertation Proposal}
\author{Kuang-Chen Lu}
\date{\today}

\bibliography{bibfile}

\usepackage{geometry} % Adjust page margins
\usepackage{amsmath, amssymb} % For math symbols and equations
\usepackage{graphicx} % For including images
\usepackage{titlesec} % For customizing section titles
\usepackage{biblatex} % For managing references
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\miscon}[1]{\textbf{#1}}
\newcommand{\surmisedMiscon}[1]{\miscon{#1}$^\ddag$}
\newcommand{\pValue}[0]{\textit{p}\nobreakdash-value}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{tabularx}   %% For tables with growable column
\usepackage{cleveref}   %% For smart ref
\usepackage{graphicx}   %% For inserting images
\graphicspath{ {./images/} }  %% configure the graphicx package
\usepackage{multirow}
\usepackage{multicol}
\usepackage{ragged2e}  %% For \raggedright
\usepackage{listings}  %% For typesetting code
\usepackage{enumitem}  %% For noitemsep and topsep
\lstdefinelanguage{smol} {
  alsoletter={ABCDEFGHIJKLMNOPQRSTUVWXYZabcdefghijklmnopqrstuvwxyz-!?*},
  morekeywords={defvar,deffun,lambda,set!,let,let*,letrec,begin,if,cond,else,and,or},
  morecomment=[l]{;},
  morestring=[b]",
}
\lstset{
  basicstyle=\small\ttfamily,
  breaklines=true,
  language=smol,
  % aboveskip=-0.3 \baselineskip,   % remove extra space before
  % belowskip=-0.8 \baselineskip,     % and after code blocks
}
%% Custom commands:
%% new terms are introduced with a boldface
%% (the common style is italic, but KC thinks bf is easier to locate)
\newcommand{\newterm}[1]{\textbf{#1}}

%% abbrev. for JavaScript etc.
\newcommand{\js}{JavaScript}
\newcommand{\py}{Python}

%% abbrev. for quizzes
\newcommand{\QuizFun}{Quiz 1}
\newcommand{\QuizState}{Quiz 2}
\newcommand{\QuizHof}{Quiz 3}
\newcommand{\quizFun}{Quiz 1}
\newcommand{\quizState}{Quiz 2}
\newcommand{\quizHof}{Quiz 3}

%% abbrev. for tutorials
\newcommand{\TutDef}{The \textbf{def} tutorial}
\newcommand{\TutSet}{The \textbf{set} tutorial}
\newcommand{\TutVec}{The \textbf{vec} tutorial}
\newcommand{\TutLam}{The \textbf{lam} tutorial}
\newcommand{\TutLet}{The \textbf{let} tutorial}
\newcommand{\tutDef}{the \textbf{def} tutorial}
\newcommand{\tutSet}{the \textbf{set} tutorial}
\newcommand{\tutVec}{the \textbf{vec} tutorial}
\newcommand{\tutLam}{the \textbf{lam} tutorial}
\newcommand{\tutLet}{the \textbf{let} tutorial}

%% abbrev. for referring to supplementary materials
\newcommand{\supplRef}[1]{\texttt{#1} in the supplementary materials}

\begin{document}

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Misconceptions about core linguistic concepts like mutable
  variables, mutable compound data, and their interaction with scope
  and higher-order functions seem to be widespread.
  But
  how do we detect
  them, given that experts have blind spots and may not realize the
  myriad ways in which students can misunderstand programs?
  Furthermore, once identified, what can we do to correct them?

  In this paper, we present a curated list of misconceptions, and an
  instrument to detect them. These are distilled from student work
  over several years and match and extend prior research.  We also
  present an automated, self-guided tutoring system. The tutor builds
  on strategies in the education literature and is explicitly designed
  around identifying and correcting misconceptions.

  We have tested the tutor in multiple settings.
  Our data consistently show that (a) the misconceptions we tackle are
  widespread, and (b) the tutor appears to improve understanding.
\end{abstract}

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
\label{s:intro}

A large number of widely used modern programming languages share a
common semantic basis:
\begin{itemize}

  \item lexical scope
  \item nested scope
  \item eager evaluation
  \item sequential evaluation (per ``thread'')
  \item mutable first-\emph{order} variables
  \item mutable first-\emph{class} structures (objects, vectors, etc.)
  \item higher-order functions that close over bindings
  \item automated memory management (e.g., garbage collection)

\end{itemize}
This semantic core can be seen in languages from ``object-oriented''
languages like C\# and Java, to ``scripting'' languages like
JavaScript, Python, and Ruby, to ``functional'' languages like the ML
and Lisp families. Of course, there are sometimes restrictions
(e.g., Java has restrictions on closures) and extensions (such as the
documented semantic oddities of JavaScript and Python \cite{bernhardtWat2012, guhaEssenceJavascript2010, politzPythonFullMonty2013, politzTestedSemanticsGetters2012}). Still, this semantic core bridges many syntaxes, and
understanding it helps when transferring knowledge from old languages
to new ones. In recognition of this deep commonality, in this paper we
choose to call this the \emph{Standard Model of Languages} (SMoL).

Unfortunately, this combination of features appears to also be non-trivial for
programmers to understand. Consider the following scenario: to create a
calculator, we have to construct a callback that can be attached to
each button. The following Python program seems to achieve the
construction of these callbacks and the act of pushing the buttons in
order:
\newpage
\begin{lstlisting}[language=Python]
button_list = []

for i in range(10):
    button = lambda: print(i)
    button_list.append(button)

for button in button_list:
    button()
\end{lstlisting}
That is, a user would expect to see \verb|0| through \verb|9|. In
fact, however, it prints \verb|9| ten times.

As the related work section (\Cref{s:rel-work}) describes,
multiple researchers, in different
countries and different kinds of post-secondary educational contexts,
have studied how students fare with scope and state. They consistently
find that even advanced students have difficulty with such programs
and even programs simpler than this.

In fact, these problems are not at all limited to students. This
specific looping problem even trips up industrial programmers. The
C\# language \emph{changed} to produce \verb|0| through \verb|9| \cite{lippertClosingLoopVariable2009}. It is now also the focus of a language design
change in Go \cite{chaseFixingLoopsGo2023} for a new kind of looping
construct \cite{coxSpecLessErrorprone2023}.
It continues to trip up programmers
(e.g., \cite{sharveyCreatingFunctionsLambdas2022})
despite being
documented as a ``gotcha'' in Python \cite{reitzHitchhikerGuidePython2016}.
% SK: But not everyone has read the document. I don't think the document is
% a Python official document. Also, this sentence sounds like stating that
% the cited Python programmers are stupid.

Most of the time, however, these misunderstandings do not lead to
language design changes. Nor are changes necessarily desirable: the
SMoL feature set has presumably evolved because it is convenient
for writing non-trivial programs (e.g., mutable structures) without
being too unwieldy (e.g., no dynamic scope). Furthermore, having a
good mental model of these features
is essential to understand ownership \cite{clarkeOwnershipTypesFlexible1998},
manage parallelism, and more. Thus, we
still need to train students and other programmers on the semantic
features \emph{and their interactions} in SMoL.

In response, we have (a) run a multi-year, multi-phase study to
identify these misunderstandings while trying to work around the blind
spots of experts (and also drawing on prior work),
and (b) constructed the SMoL Tutor, an interactive tutor to address
these misunderstandings. The SMoL Tutor is not a \emph{programming} tutor;
rather, it assumes the user has some basic facility with programming
(i.e., that they are already familiar with concepts like vectors,
mutation, and higher-order functions). It is built around SMoL's
features and their interactions, and draws on cognitive and
educational psychology concepts to expressly identify
and correct misconceptions.

Concretely, we make the following contributions:
\begin{enumerate}

  \item We provide a collection of brief programs for detecting
        misconceptions about the features of SMoL and their
        interaction.
        These programs are short and
        most are readily translatable to a wide variety of
        programming languages that share the SMoL characteristics.

      \item We show that multiple populations have problems with these
        programs.

      \item We present a list of curated misconceptions generated
        after multiple rounds of analysis.

      \item We implement these misconceptions as interpreters,
        which find weaknesses in our manual analysis and hold
        potential for generative use in the future.

  \item We present a tutoring system that measurably corrects these
        misconceptions.

\end{enumerate}

\paragraph{A Note on Terminology} We use the term
``behavior'' to refer to the meaning of programs in terms of the
answers they produce. A more standard term for this would, of course, be
``semantics''. However, the term ``semantics tutor'' might
mislead some readers into thinking it
teaches people to read or write a formal semantics, e.g., an
introduction to ``Greek'' notation. Because that is not the kind of
tutor we are describing, to avoid confusion, we use the term
``behavior'' instead.

\section{Background: Tutoring Systems and Pedagogic Techniques}
\label{s:background}

There is an extensive body of literature on tutoring systems (\cite{vanlehnBehaviorTutoringSystems2006a} is
a quality survey), and indeed whole conferences are dedicated to
them. We draw on this literature. In particular, it is common in the
literature to talk about a ``two-loop'' architecture \cite{vanlehnBehaviorTutoringSystems2006a} where the outer
loop iterates through ``tasks'' (i.e., educational activities) and the inner loop iterates through UI events within a task. We follow the same
structure in our tutor (\Cref{s:tutor-uiux}).

Many tutoring systems focus on teaching programming
(such as the well-known and heavily studied LISP Tutor \cite{andersonLISPTutor1985}), and in
the process undoubtedly address some program behavior
misconceptions. Our SMoL Tutor differs in a notable way: it does not
try to teach programming per se. Instead, it assumes a basic
programming background and focuses entirely on program behavior misconceptions
and correcting them. We are not aware of a tutoring system (in computer science) that has
this specific design.

The SMoL Tutor is firmly grounded in one technique from
cognitive and educational psychology. The fundamental problem is: how
do you fix a misconception? One approach is to only ``present the
right answer'', for fear that discussing the wrong conception might
actually reinforce it. Instead, there is a body of literature starting
from \cite{posnerTheoryConceptualChange1982} that presents a theory of conceptual
change, at whose heart is the \emph{refutation text}. A refutation
text tackles the misconception directly, discussing and providing a
refutation for the incorrect idea. Several studies \cite{schroederRefutationTextFacilitates2022} have shown
their effectiveness in multiple other domains.

The SMoL Tutor's content structure is also influenced by work on \emph{case comparisons}
(which draws analogies between examples).
\cite{alfieriLearningCaseComparisons2013a} suggests that asking
(rather than not asking) students to find similarities between cases,
and providing principles \emph{after} the comparisons (rather than
before or not at all), are associated with better learning
outcomes.

\section{The SMoL Language}\label{s:smol}

\begin{figure}[t]
  \centering
  \begin{lstlisting}
    t ::= d
        | e
    d ::= (defvar x e)
        | (deffun (f x ...) body)
    e ::= c
        | x
        | (lambda (x ...) body)
        | (let ([x e] ...) body)
        | (begin e ... e)
        | (set! x e)
        | (if e e e)
        | (cond [e body] ... [else body])
        | (cond [e body] ...)
        | (e e ...)
    body    ::= t ... e
    program ::= t ...
  \end{lstlisting}
  \caption{The syntax of SMoL.}
  \label{f:smol-syntax}
\end{figure}

\begin{table}[t]
  \centering\begin{tabularx}{\linewidth}{r X}
    \textbf{Operators}            & \textbf{Meaning}                                       \\\hline
    \texttt{+ - * /}              & Arithmetic Operators                                   \\
    \texttt{< > <= >=}            & Number comparison                                      \\
    \texttt{mvec}                 & Create a (mutable) vector (a.k.a. array)               \\
    \texttt{vec-ref}              & Look up a vector element                               \\
    \texttt{vec-set!}             & Replace a vector element                               \\
    \texttt{vec-len}              & Get the length of a vector                             \\
    \texttt{mpair}                & Create a 2-element (mutable) vector                              \\
    \texttt{left right}           & Look up the first/second element of a 2-element vector \\
    \texttt{set-left! set-right!} & Replace the first/second element of a 2-element vector \\
    \texttt{eq?}                    & Equality                                               \\
  \end{tabularx}
  \caption{Primitive operators of SMoL.}
  \label{t:smol-primitives}
\end{table}

SMoL is designed to capture common features of many modern
languages.
The syntax of SMoL is presented in~\Cref{f:smol-syntax}, where
\lstinline|t| stands for terms, \lstinline|d| stands for definitions, \lstinline|e|
stands for expressions, \lstinline|c| stands for constants (i.e., number,
boolean, and string), and \lstinline|x| and \lstinline|f| are identifiers
(variables). The last kind of expression (i.e., \lstinline|(e e ...)|) is
function application.

SMoL's Lispy syntax is
an artifact of the course (\Cref{s:univ-1-course}) using Racket
\cite{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
however, it also proved to be pedagogically valuable.
We have found the parentheses useful when discussing
scope in conjunction with local-binding features like \lstinline|let|.
This avoids the various confusing ``variable lifting'' semantics found
in languages like Python (\cite{politzPythonFullMonty2013}; see also
posts like \cite{froadieWhatScopeVariable2022}), where the actual defined range of a
variable is not apparent from the source code.

Nevertheless, most SMoL programs are easy to translate to other
languages. We present machine-translated Python and JavaScript
versions of our programs in \supplRef{Translated\_Programs.html}. Furthermore, we intend to
make a multi-lingual tutor in the future (\Cref{s:discussion}).

The semantics of SMoL is as described in \Cref{s:intro}.
SMoL includes limited primitive operators (\Cref{t:smol-primitives}) to work with numbers, strings, and
vectors. It provides only one equality operator, which tests for exact
equality between atomic values and for pointer equality between other
values.

Students are given a working implementation of SMoL, built
as a \verb|#lang| language inside Racket \cite{felleisenProgrammableProgrammingLanguage2018}. This language
provides only the defined syntax and semantics of SMoL, with no
(other) Racket features present.
(As in Racket, arguments evaluate left-to-right, to give stateful programs an
unambiguous semantics.) The implementation is available online
[URL anonymized].

\section{Paper Roadmap}
\label{s:roadmap}
\label{s:univ-1-course}

This paper describes a four-year effort to obtain a quality instrument
to measure misconceptions and to build a tutor to address them.

The first step was a two-year process to generate (a) programs that
tend to trip up students and (b) incorrect responses to those programs
(\Cref{s:gen-prob-quizius}), using a tool for the purpose
(\Cref{s:quizius}). These were then curated (\Cref{s:smol-quizzes}) to
produce an instrument that was tested for one more year
(\Cref{s:misconceptions-quiz}). This underwent one more round of
curation to produce the final instrument~(\Cref{s:tutor-refine-program-set}).

We then present the tutor (\Cref{s:system-tutor}). In particular, we
present the list of misconceptions---grounded in student data---that
are the focus of this project (\Cref{s:misconceptions-tutor}). We also
evaluate the tutor's effectiveness at correcting them
(\Cref{s:effect-tutor}).

All this work is done with students in a ``Principles of Programming
Languages'' class at a selective, private US university. The class has
about 70--75 students per year. It is not required, so students take
it by choice. Virtually all are computer science majors. Most are in
their third or fourth year of tertiary education; about 10\% are
graduate students. All have had at least one semester of imperative
programming, and most have significantly more experience with it. Most
have had close to a semester of functional programming. The student
work described here was required, but students were graded on effort,
not correctness.

Naturally, we should wonder to what extent the demographic affects the
results: we may simply be studying weaknesses at this institution! We
therefore work with two other populations
(\Cref{s:generality-tutor}). In both, we find similar results.

\section{Generating and Collating Problems}
\label{s:gen-prob-quizius}

In \Cref{s:rel-work}, we discuss several papers that have provided
reports of student misconceptions with different fragments of
SMoL. However, it is difficult to know how comprehensive these
are. While some are unclear on the origin of their programs, they
generally seem to be expert-generated.

The problem with expert-generated lists is that they can be quite
incomplete. Education researchers have documented the
phenomenon of the \emph{expert blind spot} \cite{nathanExpertBlindSpot2001}: experts
simply do not conceive of many learner difficulties.
Thus, we need
methods to identify problems beyond what experts conceive.

Additionally, in this paper, we intentionally use the word
mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
for any reason (e.g., selecting the wrong answer from a menu). A
misconception, however, implies a conceptual problem: the student has
formed an incorrect concept in their head. For instance, they may
think that mutable structures are copied on function calls, or that
scope is resolved dynamically. This requires probing what they are
thinking.

Finally, we are inspired by the significant body of education research
on \emph{concept inventories} \cite{hestenesForceConceptInventory1992} (with a growing
number for computer science, as a survey lists \cite{taylorComputerScienceConcept2014}). In
terms of mechanics, a concept inventory is just an instrument
consisting of multiple-choice questions (MCQs), where each question has
one correct answer and several wrong ones. However, the wrong ones are
chosen with great care. Each one has been validated so that if a
student picks it, we can quite unambiguously determine \emph{what
  misconception the student has}. For instance, if the question is
``What is \verb|sqrt(4)|?'', then \verb|37| is probably an
uninteresting wrong answer, but if people appear to confuse
square-roots with squares, then \verb|16| would be present as an
answer.\footnote{Concept inventories are thus useful in many
  settings. For instance, an educator can use them with clickers to
  get quick feedback from a class. If several students pick a
  specific wrong answer, the educator not only knows they are wrong,
  but also has a strong inkling of \emph{precisely what} misconception
  that group has and can address it directly. We expect our
  instruments to be useful in the same way.}

All these, however, add up to a somewhat challenging demand. We want
to produce a list of questions (each one an MCQ) such that
\begin{enumerate}

\item we can get past the expert blind spot,

\item we have a sense of what misconceptions students have, and

\item we can generally associate wrong answers with specific
  misconceptions, approaching a concept inventory.

\end{enumerate}

\subsection{Generating Problems Using Quizius}
\label{s:quizius}

Our main solution to the expert blind spot is to use the Quizius system
\cite{saarinenHarnessingWisdomClasses2019}. In contrast to the very
heavyweight process (involving a lot of expert time)
that is generally
used to create a concept inventory, Quizius uses a lightweight,
interactive approach to obtain fairly comparable data, which an expert
can then shape into a quality instrument.

In Quizius, experts create a prompt; in our case, we asked students to
create small but ``interesting'' programs using the SMoL
language. Quizius shows this prompt to students and gathers their
answers. Each student is then shown a set of programs created by other
students and asked to predict (without running it) the value produced
by the program.\footnote{In the course (\Cref{s:univ-1-course}),
  students were given credit for using Quizius but not penalized for wrong
  answers, reducing their incentive to ``cheat'' by running
  programs. They were also told that doing so would diminish the value of
  their answers. Some students seemed to do so anyway, but most
  honored the directive.} Students are also asked to provide a rationale for why
they think it will produce that output.

Quizius runs interactively during an assignment period.
At each point, it needs to determine which
previously authored program to show a student.
It can
either ``exploit'' a given program that already has responses
or ``explore'' a new
one. Quizius thus treats this as a multi-armed bandit problem
\cite{katehakisMultiArmedBanditProblem1987} and uses that to choose a program.

The output from Quizius is (a) a collection of programs; (b) for
each program, a collection of predicted answers; and (c) for each
answer, a rationale. Clustering the answers is easy (after
ignoring some small syntactic differences). Thus, for
each cluster, we obtain a set of rationales.

After running Quizius in the course (\Cref{s:univ-1-course}),
we took over as experts. Determining which is the right
answer is easy. Where expert knowledge is useful is in
\emph{clustering the rationales}. If all the rationales for a wrong
answer are fairly similar, this is strong evidence that there is a
common misconception that generates it. If, however, there are
multiple rationale clusters, that means the program
is not discriminative enough to distinguish the misconceptions, and it
needs to be further refined to tell them apart. Interestingly, even
the correct answer needs to be analyzed, because sometimes correct
answers do have incorrect rationales (again, suggesting the program
needs refinement to discriminate correct conceptions
from misconceptions).

Prior work using Quizius \cite{saarinenHarnessingWisdomClasses2019} finds that students do author programs
that the experts did not imagine. In our case, we seeded Quizius with
programs from prior papers (\Cref{s:rel-work}), which gives the first
few students programs to respond to. However, we found
that Quizius significantly expanded the scope of our problems and
misconceptions. In our final instrument, most programs were directly
or indirectly inspired by the output of Quizius.

\subsection{Collating Problems}
\label{s:smol-quizzes}

While Quizius is very useful in principle, it also produced data that
needed significant curation for the following reasons:
\begin{itemize}

\item A problem may have produced diverse outputs simply because it
  was written in a very confusing way. Such programs do not reveal
  any useful \emph{behavior} misconceptions, and must therefore be
  filtered out. For instance:

  \begin{lstlisting}
  (defvar x 1)
  (defvar y 2)
  (defvar z 3)
  (deffun (sum a ...) (+ a ...))
  (sum x y z)
  \end{lstlisting}

  A reader might think that \verb|sum| takes variable arguments (so
  the program produces \verb|6|), but in fact \verb|...| is a single
  variable, so this produces an arity error.

\item Some programs relied on (or stumbled upon) underspecified aspects
  of unimportant (in particular, non-standard) parts of SMoL, such as floating-point versus
  rational arithmetic.

\item A problem may have produced diverse outputs simply because it is
  hard to parse or to (mentally) trace its execution. One example was a
  17-line program with 6 similar-looking and -named functions. As
  another example:

\begin{lstlisting}
  (defvar a (or (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
  (defvar b (and (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
  (defvar c (and (/ 1 0.0) (/ 1 (- 0.25 -0.25))))
  (defvar d (or (/ 1 0) (/ 1 (- 0.25 -0.25))))
  (and (or a c) (or b d))
\end{lstlisting}

  This program is not only confusing, it \emph{also} tests interpretations of (a) exact
  versus inexact numbers and (b) truthy/falsiness, leading to
  significant (but not very useful) answer diversity.

\item As noted above, a program's wrong (or even correct) answers may
  correspond to multiple (mis)conceptions. In these cases, the program
  must be refined to be more discriminative.

\end{itemize}
and so on.
We therefore manually curated the Quizius output
to address these issues.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{images/smol-quizzes-example.png}
  \caption{Screenshot of a SMoL Quiz question.}
  \label{f:smol-quizzes-example}
\end{figure}

\subsection{The SMoL Quizzes}%
\label{s:misconceptions-quiz}

Having curated the output, we had to confirm that these programs were
still effective! That is, they needed to actually find student errors.

We therefore turned the programs into a set of quizzes (in the US
sense: namely, a brief test of knowledge) that we call the SMoL
Quizzes. There were three quizzes, ordered by linguistic
complexity. The first consisted of only basic operators and
first-order functions. The second added variable and structure
mutation. The third added \lstinline|lambda| and higher-order functions.

The goal of the SMoL Quizzes was to confirm that the aforementioned
processes of cleansing and enriching the problems was successful. The
quizzes were therefore administered in the third year of this project in the
same course.
\Cref{f:smol-quizzes-example} shows a
sample question. Every question got an ``Other'' option. If chosen,
the quiz gave the user a text box with the caption ``Please
specify''. The goal here was to record any other answers, which in
turn might lead to fresh misconceptions.

Question orders were partially randomized. We wanted students to get
some easy, warm-up questions initially, so those were kept at the
beginning. Similarly, we wanted programs that are syntactically
similar to stay close to each other in the quiz. This is so that, when
students got a second such program, they would not have to look far to
find the first one and confirm that they are indeed (slightly)
different, rather than wonder if they were seeing the same program
again.

Students only received feedback after having completed a whole
quiz. At the end of each quiz, they received both summative feedback
\emph{and} a refutation text that explained every program that
appeared in the quiz. Students were also encouraged to run the
programs, but we have little reason to believe that they did (and
certainly they asked few questions on the class forum about them).

Due to space limitations, we present the entire instrument in \supplRef{SMoL Quizzes/instrument}.
Here we focus on a few programs where student choices correspond to
misconceptions identified earlier, thereby also showing
that the curated programs are
still effective.
Syntactically, \verb|#t| and \verb|#f| are true and false,
while \verb|#(...)| is a vector. We use a * to indicate the correct
answer (which also matches the implementation's output).

\subsubsection{Quiz 1: Basic Operators and First-Order Functions}%
\label{s:misconceptions-quiz:fun}

% This quiz starts with two warm-up and not very interesting questions
% and then other questions. The order of the other questions are fully
% randomized.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l l}
    \textbf{Questions}                       &
    \textbf{Frequency Table of Answers}          \\\hline
    \lstinputlisting{programs/def_twice.txt} & {
        \begin{tabularx}{1in}{r l}
          $^\dag$59\%  & \verb|2 0|             \\
          *32\% & \verb|Error|           \\
          4\%   & \verb|Other: 2 2|      \\
          3\%   & ``Nothing is printed'' \\
          2\%   & \verb|0 0|             \\
        \end{tabularx}
    }                                            \\\hline
    \lstinputlisting{programs/dyn_scope.txt} & {
        \begin{tabularx}{1in}{r l}
          *65\% & \verb|Error| \\
          $^\dag$29\%  & \verb|42 #t| \\
          6\%   & Other        \\
        \end{tabularx}
    }                                            \\\hline
    \lstinputlisting{programs/reassign.txt}  & {
        \begin{tabularx}{1in}{r l}
          *53\% & \verb|#(6 5) 5|                 \\
          *20\% & \verb|Error|                    \\
          $^\dag$11\%  & \verb|#(6 6) 5|                 \\
          $^\dag$11\%  & \verb|#(6 6) 6|                 \\
          5\%   & ``Nothing is printed'' or Other \\
        \end{tabularx}
    }                                            \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizFun. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}. Wrong answers that we discuss are
    marked with $^\dag$.}
  \label{t:quizzes-data-fun}
\end{table}
\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that we
consider the most interesting. These data confirm the presence of
scope-related misconceptions:
\begin{enumerate}
  \item At least 59\% of students incorrectly believe that a
        variable can be defined twice in one block.
  \item 29\% of students believe that a variable defined in a function
        will be available in the top-level (or global) environment
        after the function is called. That is, these students may have a
        dynamic scope misconception.
  \item 22\% (11\% + 11\%) of students believe variables themselves
        can be passed as arguments and redefined inside the function.
        They disagree on whether the redefinition persists after the
        function call.
\end{enumerate}

\subsubsection{Quiz 2: Adding Variable and Structure Mutation}%
\label{s:misconceptions-quiz:state}

% This quiz randomizes the order of all questions.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l l}
    \textbf{Question}                           &
    \textbf{Frequency Table of Answers}             \\\hline
    \lstinputlisting{programs/circularity.txt}  & {
        \begin{tabularx}{1in}{r p{2.5in}}
          $^\dag$50\%  & \verb|#(#(2 #(2 3)) #(2 3))| \\
          *29\% & \verb|x=#(x x)| or
          something similar. Both \verb|(left x)| and \verb|(right x)|
          are \verb|x| itself.                 \\
          $^\dag$16\%  & \verb|Error|                 \\
          5\%   & Other                        \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/mvec_in_mvec.txt} & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *62\% & \verb|#(#(100 2 3 4) #(100 2 3 4))| \\
          $^\dag$26\%  & \verb|#(#(1 2 3 4) #(100 2 3 4))|   \\
          6\%   & \texttt{Error}                      \\
          4\%   & \verb|#(#(1 2 3 4) #(1 2 3 4))|     \\
          2\%   & Other                               \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/vset_in_let.txt}  & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *59\% & \verb|#(10)|  \\
          $^\dag$36\%  & \verb|#(123)| \\
          5\%   & Other         \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/var_as_arg.txt}   & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *65\% & \texttt{12} \\
          $^\dag$31\%  & \texttt{0}  \\
          4\%   & Other       \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/setset.txt}       & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *59\% & Other: \texttt{5 7}   \\
          $^\dag$27\% & \texttt{6 7}          \\
          11\%  & \texttt{5 5}          \\
          2\%  & Other: \texttt{Error} \\
          1\%  & Other: \texttt{5 6}   \\
        \end{tabularx}
    }                                               \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizState. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}. Wrong answers that we discuss are
    marked with $^\dag$.}
  \label{t:quizzes-data-state}
\end{table}
\Cref{t:quizzes-data-state} lists interesting programs after the
addition of state. These data suggest the students have aliasing-related misconceptions:
\begin{itemize}
  \item Up to 50\% of students think vectors are copied rather than
  aliased.
  \item 16\% of students think trying to construct and print a
  self-referring vector would error. (This program is ambiguous:
  \emph{constructing} works in most SMoL languages, but \emph{printing}
  may well cause a problem. These identified ambiguities are
  addressed in \Cref{s:tutor-refine-program-set}.)
  \item 31\% of students think a variable is aliased by a parameter if
        the two variables have the same name. Perhaps interestingly,
        fewer students (27\%) think the variable aliasing would
        happen if the two variables had different names.
\end{itemize}

\subsubsection{Quiz 3: Adding Closures and Higher-Order Functions}%
\label{s:misconceptions-quiz:hof}

% This quiz includes two groups of questions. Each group contains
% programs that are really similar to each other. The order of the two
% groups and the other questions are randomized. Question order within
% each group is also randomized. The first two programs in the
% \Cref{t:quizzes-data-hof} belongs to one group, which is about
% changing a function's behavior by mutating a variable defined outside
% the function. The other group is about equality between functions. The
% latter group is not included in the main text because we think it is
% not very related to the theme of this paper.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l l}
    \textbf{Question}                                 &
    \textbf{Frequency Table of Answers}                   \\\hline
    \lstinputlisting{programs/fun_returns_lambda.txt} & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *82\% & \verb|3|     \\
          $^\dag$17\%  & \verb|Error| \\
          1\%   & Other        \\
        \end{tabularx}
    }                                                     \\\hline
    \lstinputlisting{programs/fun_and_state_1.txt}    & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *74\% & \verb|4|            \\
          $^\dag$21\%  & \verb|3|            \\
          6\%   & Other: \verb|Error| \\
        \end{tabularx}
    }                                                     \\\hline
    \lstinputlisting{programs/fun_and_state_2.txt}    & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *88\% & \verb|4|            \\
          $^\dag$10\%  & \verb|3|            \\
          3\%   & Other: \verb|Error| \\
        \end{tabularx}
    }                                                     \\\hline
    \lstinputlisting{programs/counter.txt}            & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *62\% & \verb|1 1 2 3 2| \\
          $^\dag$34\%  & \verb|1 1 1 1 1| \\
          4\%   & Other            \\
          1\%   & \verb|1 1 2 3 4| \\
        \end{tabularx}
    }                                                     \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizHof. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}. Wrong answers that we discuss are
    marked with $^\dag$.}
  \label{t:quizzes-data-hof}
\end{table}
\Cref{t:quizzes-data-hof} lists interesting programs
after the addition of closures and higher-order functions.
It extends what we saw with loops in
\Cref{s:intro}: that students have misconceptions about their interaction with mutable variables:
\begin{itemize}
  \item 17\% of students think \texttt{lambda} functions can't refer
        to free variables.
  \item 21\% of the students think mutating a variable defined outside
        a lambda can't possibly change the behavior of the lambda.
        Perhaps interestingly, this
        misconception seems to depend on how the lambda is constructed
        (compare the middle two programs).
  \item Student understanding of the
        \lstinline|let|-over-\lstinline|lambda| pattern is weak.
\end{itemize}

\section{The SMoL Tutor}
\label{s:system-tutor}

So far, we have focused on identifying problems. As noted earlier
(\Cref{s:smol-quizzes}), we provided students with refutation texts after the quizzes,
but it is unclear to what extent students read, understood, or
internalized these. We also wanted to determine whether other
populations run into these issues, and it was unclear whether we would
be able to administer passive quizzes to them.

Furthermore, while the quizzes may be a useful diagnostic, our goal is
not only to find faults but to improve the understanding of basic
programming language behavior. We believe (and presumably so does
anyone else who writes a formal semantics!)\ that an understanding of
these basic program behaviors is important, and even more so when it
comes to understanding concurrency, ownership, and other advanced
features. Even more simply, misunderstanding these clearly impacts
development and debugging time.

In response, we decided to create a \emph{tutor}: the SMoL Tutor. It is built
around our quiz instruments and the detected misconceptions. We
describe the Tutor from a user's perspective in \Cref{s:tutor-uiux}, and
explain how it was populated from the
SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
We then discuss what
we learned from its data (\Cref{s:misconceptions-tutor}), and finally
evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).

\subsection{The User Experience}
\label{s:tutor-uiux}

The Tutor covers five major topics, shown on the left in
\Cref{t:tutor-goal-sentences}. The larger topics are further broken
down into 2--3 modules. The goal was for students to spend at most
20--30 minutes per module. Our data show that in practice, students
spent about 9.8 (median) minutes.

Each tutorial consists of a sequence of
questions. Most questions in the Tutor are \newterm{interpreting
questions}.\footnote{The Tutor also asked students to perform some
other activities, like showing programs and asking for their heap
content, which we do not cover in this paper.}
 These questions (illustrated in
\Cref{f:smol-tutor-example}) are versions of the SMoL
Quizzes (obtained by the process described in \Cref{s:tutor-refine-program-set}).
In each of these questions, students are shown a program
and asked to predict the program's running result(s) by answering an
MCQ (with an ``Other'' option). After making a choice, students receive
feedback. If a student chooses incorrectly, they are (a) given an
explanation that is \emph{based on the misconception associated with
  that wrong answer} (or a generic one, if there is not a specific
misconception),
and then (b) asked to answer a
second question:
\begin{enumerate}

\item The second question is semantically the same as the first, but
  with superficial changes (e.g., variable names, constants, and operators are
  changed) so that the student cannot immediately
  guess the answer.

\item Instead of multiple-choice, students must \emph{type} the answer
  into a text box. (The Tutor normalizes text to accommodate variations.)
  This is
  intentional. First, we wanted to force reflection on the
  explanation just given, whereas with an
  MCQ, students could just guess. Second, we felt that students would
  find typing more onerous than clicking. In case students
  had just guessed on a question, our hope is that the penalty of having to
  type means, on \emph{subsequent} tasks, they
  would be more likely to pause and reflect before choosing.

\end{enumerate}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{images/smol-tutor-example.png}
  \caption{Screenshots of an interpreting question in SMoL Tutor. The
    top-left shows the initial state, where the question is presented
    as an MCQ. If a student chooses a wrong answer, they will receive
    feedback (bottom-left) and will be asked a similar question
    (right). The similar question must be answered by typing.}
  \label{f:smol-tutor-example}
\end{figure}

In addition to asking students questions, the Tutor along the way
introduces terminology and states the true conceptions. These are the
teaching goals of the Tutor. We therefore refer to these as
\newterm{goal sentences}. \Cref{t:tutor-goal-sentences} lists
the (abbreviated) goal sentences for each tutorial.
Readers can find the full Tutor in \supplRef{SMoL Tutor/instrument}.

\begin{table}[t]
  % \begin{tabularx}{250pt}{|c|X|c|X|} \end{tabularx}
  \centering\begin{tabularx}{\linewidth}{
      >{\hsize=.5\hsize\linewidth=\hsize}X
      >{\hsize=1.5\hsize\linewidth=\hsize}X}
    \textbf{Tutorial}                             & \textbf{Goal Sentences}
    \\\hline
    \newterm{scope}: Variable definitions and function definitions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Referring to an unbound variable leads to an error.
      \item Define "blocks".
      \item SMoL is lexically scoped rather than dynamically scoped.
      \item SMoL disallows defining a variable twice in one block.
      \item SMoL is eager (and evaluates from left to right) rather
            than lazy, reactive, or relational.
    \end{enumerate}
    \\\hline
    \newterm{mut-vars}: Variable updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Variable assignment respects the hierarchical structure of blocks.
      \item Variables do not alias.
    \end{enumerate}
    \\\hline
    \newterm{begin}: Sequencing expressions &
    A sequencing expression evaluates its sub-expressions from left to right and
    returns the value of the last sub-expression.
    \\\hline
    \newterm{vectors}: Vectors and vector updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Define heap and memory addresses.
      \item Vectors can be referred to by vectors, including themselves.
      \item Vectors are not copied but aliased by bindings.
      \item Constructing vectors doesn't alter environments.
      \item Introducing bindings doesn't alter the (value part of) heap.
    \end{enumerate}
    \\\hline
    \newterm{lambda}: Lambda expressions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Functions are (first-class) values.
      \item Functions remember the environment where they were created.
      \item Functions can be created with lambda expressions.
      \item A function definition can be viewed as a variable definition plus a lambda expression.
    \end{enumerate}
    \\\hline
    \newterm{local}: Local binding forms &
    Introduce \texttt{let}, \texttt{letrec}, and \texttt{let*}.
    % \\\hline
  \end{tabularx}
  \caption{SMoL Tutor tutorials and their goal sentences.}
  \label{t:tutor-goal-sentences}
\end{table}

Some later tutorials include questions about earlier tutorials so
that we can check whether students remember the concepts across
modules. In particular, the \textbf{mut-vars} tutorial starts with questions
about \textbf{scope}, and \textbf{lambda} starts with questions about
\textbf{mut-vars} and \textbf{vectors}.

Of the modules, \textbf{local} is the least portable across
languages. While local binding is of course present in other
languages, this is mostly covered using nested \lstinline|defvar|s in
earlier modules, starting with \textbf{scope}. The distinction central
to this module (between three local-binding constructs with slightly
different scopes) is primarily a focus of Lispy languages. Therefore,
we exclude this module from our analysis and instruments, and indeed
plan to deprecate the module in future versions.

\Cref{s:smol-quizzes} discusses the partial randomization employed by
the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
randomize question order. This is because the Tutor consists of more
than just questions: it also has explanatory text. This text is based
on the preceding problems. Authoring it is therefore somewhat like
writing a textbook, with complex dependencies that can easily be
broken.

\subsection{Collating Problems for the Tutor}%
\label{s:tutor-refine-program-set}

The SMoL Tutor's interpreting questions are the final instrument of
this paper.
We bootstrap it from the SMoL
Quizzes, but made the following alterations:
\begin{enumerate}

  % fill gaps in goal sentences
  \item In the process of developing the goal sentences, we felt the
        existing programs were insufficient to cover the ideas we wanted
        students to work through. For instance, SMoL Quizzes did not
        alias vectors using \lstinline|defvar|, so we added this program:

\begin{lstlisting}
(defvar x (mvec 100))
(defvar y x)
(vec-set! x 0 200)
y
\end{lstlisting}

  \item We renamed vector operators to avoid confusion between
        \lstinline|vset!|, which replaces a vector element, and \lstinline|set!|,
        which mutates variables. We rename all vector operators from, for example,
        \lstinline|vset!| to \lstinline|vec-set!|.

  \item We deferred the local binding forms to the end, since we felt
        they were less essential. We therefore rewrote programs that use
        it to not depend on it.

        % \item We removed some questions that we deemed ``Lispy'' (e.g.,
        %   involving the difference between parenthetical and infix syntax).

  \item We removed some programs that relied on underspecified aspects or
        unimportant aspects of SMoL. For example, some depended on
        whether operations on pairs could be applied to arbitrary vectors:

\begin{lstlisting}
  (pair? (mpair 1 2))
  (pair? (mvec 1 2))
  (pair? '#(1 2))
  (pair? '(1 2))
\end{lstlisting}

        As another example, one hinged on
        whether or not a function's formal parameter could have the same
        name as the function itself:

\begin{lstlisting}
  (deffun (f f) f)
  (f 5)
\end{lstlisting}

  \item We resolved ambiguities in some programs, either adding
    answer choices or even adding other questions to tease
    out different interpretations.

  \item To reduce the number of concepts, we removed programs that relied
        upon \emph{im}mutable vectors and lists, because they did not
        seem to create problems. (For brevity, we leave these out of
        the presentation of SMoL in \Cref{s:smol}, though they are in the implementation.)

  \item We removed questions related to function equality, which was
        not a focus of the Tutor.\footnote{However, we believe there is
          a good deal of confusion about notions of equality. We
          intend to
          add that as a major tutorial component in future versions.}

  \item We removed programs of a ``Lispy'' nature, such as one where
    the answer depended on whether the reader correctly understood
    this inequality:

\begin{lstlisting}
      (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
\end{lstlisting}

    This checks whether $\verb|n| < \verb|3|$, but some presumably
    vocalized it as ``greater than \verb|3|, \verb|n|?''.

\end{enumerate}
Most of these steps either modify or elide programs. Two cases,
filling gaps in light of the goal sentences and resolving ambiguities,
introduce
programs. Starting with the 37 programs in the SMoL Quizzes, we added
52 more programs to arrive at a total of 89.

In addition to generating this set of programs, we also
modified the answer options. In addition to retaining the correct and
incorrect answers from the SMoL Quizzes (and constructing our best
guess of analogous incorrect answers for the new problems), we added
more wrong answers.

The reason is as follows. The SMoL Quizzes often have very few wrong
choices: of the 37 tasks, 26 have only three choices (including the
correct answer and the ``Other'' option\footnote{In a few cases, we
  removed the \emph{correct} answer from the choices. A student would
  therefore have to click on ``Other'' and type it in. The idea was to
  make the ``Other'' option more salient and plausible.}),
seven have 4, and only four have 5 or 6.
Thus, in most
cases, students have a 25\% or even 33\% chance of just guessing the
right answer or successfully using a process of elimination. By
increasing the number of options, we hoped to
greatly reduce the odds of getting the right answer by chance or
by elimination.

It was important to add wrong answers that are not utterly
implausible, because those would become easy to eliminate. Therefore,
we added numeric constants mentioned in the problem, permuted some of
the values in case of multiple outputs, and so on. In general, we
ended up increasing the number of choices substantially:
only 8 out of the 89 questions have three or four choices; 70
questions have 5--8 choices; and 11 questions have 9--14 choices. We hope
this reduced both guessing and elimination, and forced students to
actually think through the program. Of course, these new answers do
not have a clear associated misconception, so their mistakes are given
the generic explanation.

\section{Misconceptions Detected by the SMoL Tutor}%
\label{s:misconceptions-tutor}

We now examine what we learned from the interpreting
questions in terms of program behavior
misconceptions. But first, we explain how we associate incorrect
answers with misconceptions.

We do not present the results of \textbf{begin} and
\textbf{local} (\Cref{t:tutor-goal-sentences}), because they focus on constructs not usually found in
non-Lispy languages. They were included in the Tutor to help students
write programs for the course, but are not interesting from a SMoL
perspective.

\subsection{Misinterpreters}
\label{s:misinterp-intro}

In principle, an expert can identify what misconceptions a particular
incorrect answer might correspond to. In practice, we found this
rather difficult for three reasons. First, with 89 programs, each with
several answers, it's easy to make mistakes. Second, our own expert blind
spot may prevent us from seeing an interpretation that would
lead to an additional association. Finally, and specific to our case,
since we had either curated or written all the programs and answers,
we were likely to miss some associations we had not intended.

To address this problem, we formalized misconceptions as
interpreters. That is, for each misconception we created a
corresponding \emph{misinterpreter}. This is an interpreter for the
SMoL syntax that intentionally has a semantic error corresponding to
that misconception. Put differently, a misinterpreter is a like a
``definitional interpreter for a misconception''.

By running programs through the misinterpreters, we can more uniformly
and rigorously identify \emph{all} the misconceptions associated with
a wrong answer. Furthermore, if we identify a new misconception (or
alter a misinterpreter), it is easy to automatically re-classify all
the answers. By using misinterpreters, we indeed found new
interpretations for existing program-answer pairs.
We provide all the misinterpreters in the artifact.

\subsection{A Catalog of Misconceptions}

We iteratively created
our final catalog of misconceptions
(from the perspective of the data in this paper).
We started
with misinterpreters representing the misconceptions described in
\Cref{s:misconceptions-quiz}. These are misconceptions for which we
have reasonable validation (due to the prose in Quizius), so we call
them \emph{grounded} misconceptions. We then looked at wrong answers
not covered by these but chosen by students, and did our best to
distill these into misconceptions. These are \emph{surmised}
misconceptions (which we
identify with a $^\ddag$), which need to be validated in the future.\footnote{To keep
  the time spent in each module reasonable, the SMoL Tutor does not
  ask students for rationales for their programs. However, we could
  easily modify it to do so a small number of times per user, to make
  progress towards this need.} We then re-ran the misinterpreters
against the chosen answers. We terminated when all the remaining wrong
answers were either (a) found in very few students (we found a gap
between 23\% and 13\%, and hence took 14\% as the threshold),
(b) difficult for
us to attribute to a misconception, or (c) appeared to us to be
``Lispy'' and hence not of broad interest.

\Cref{t:tutor-grounded-misconceptions-1,t:tutor-grounded-misconceptions-2,t:tutor-surmised-misconceptions}
list the final catalog.
For each, we also
present a Tutor question
for which the marked wrong answer can be explained by \emph{only} the
named misconception.
That is, that program-answer pair is a representative example of that
misconception.\footnote{An astute reader may well imagine another
  misinterpretation; but we presumably did not find evidence of it in our
  data. We do welcome other misinterpretations, for which we
  can add more misinterpreters!}

  \begin{table}[t]
    \centering
    \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
      \textbf{Misconception}                                                     &
      \textbf{Question}                                                          &
      \textbf{Table of Answers}
      \\\hline
      \miscon{CallByRef} Function calls alias variables.                         &
       % mut-vars.not_aliased_by_funarg_2 (35)
       \lstinputlisting{programs/call-by-ref.txt}                                                                    & {
        \begin{tabularx}{1in}{r l}
          78\%  & \verb|12|      \\
          11\% & \verb|error|      \\
          **10\%   & \verb|0|  \\
          1\%   & \verb|23|  \\
        \end{tabularx}
    }
      \\\hline
      \miscon{CallsCopyStructs} Function calls copy data structures.      &
      % vectors2.alias_with_funcall (48)
      \lstinputlisting{programs/calls-copy-structs.txt}
      & {
        \begin{tabularx}{1in}{r l}
          90\%  & \verb|#(173 0)|      \\
          **10\%   & \verb|#(1 0)|  \\
        \end{tabularx}
    } \\\hline
      \miscon{DeepClosure} Closures copy the \textit{values} of free variables.  &
      % lambda3.lambda_remembers_env (70)
      \lstinputlisting{programs/deep-closure.txt}                                & {
          \begin{tabularx}{1in}{r l}
            86\%  & \verb|4|      \\
            **10\% & \verb|3|      \\
            3\%   & \verb|error|  \\
            1\%   & \verb|lambda| \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{DefByRef} Variable definitions alias variables.                    &
      % mut-vars.not_aliased_by_defvar_1 (29)
      \lstinputlisting{programs/def-by-ref.txt}                                  & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            85\%   & \verb|0 12|                \\
            **12\% & \verb|0 0|                 \\
            2\%    & \verb|error|               \\
            1\%    & depends on implementation. \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefOrSet}  Both definitions and variable assignments are
      interpreted as follows: if a variable is not defined in the
      current environment, it is defined. Otherwise, it is mutated to
      the new value.
      &
      % program: mut-vars.update_undefined (28)
      \lstinputlisting{programs/def-or-set.txt}                                  & {
          \begin{tabularx}{1in}{r l}
            85\%   & \verb|error| \\
            **15\% & \verb|2|     \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefsCopyStructs} Variable definitions copy structures recursively. &
      % vectors2.alias_with_defvar (47)
      \lstinputlisting{programs/defs-copy-structs.txt}                                & {
          \begin{tabularx}{1in}{r l}
            **67\% & \verb|#(100)| \\
            30\%   & \verb|#(200)| \\
            1\%    & \verb|#(300)| \\
            1\%    & \verb|error|  \\
          \end{tabularx}
      }                                                                                   \\\hline
    \end{tabularx}
    \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
      with ``**'' represent the misconception. (Part I)}
    \label{t:tutor-grounded-misconceptions-1}
  \end{table}

  \begin{table}[t]
    \centering
    \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
      \textbf{Misconception}                                                     &
      \textbf{Question}                                                          &
      \textbf{Table of Answers}
      \\\hline
      \miscon{FlatEnv} There is only one environment, the global
      environment. (This misconception is a kind of dynamic scope.)              &
      % scope.error_when_refer_to_undefined (9)
      \lstinputlisting{programs/flat-env.txt}                                    & {
          \begin{tabularx}{1in}{r l}
            96\%  & \verb|error| \\
            **4\% & \verb|402|   \\
          \end{tabularx}
        }
      \\\hline
      \miscon{FunNotVal} Functions are \textit{not} considered
      first-class values. They can't be bound to other variables, passed as arguments, or referred to by data structures.
                                                                                 &
      % lambda1.fun_as_parameter (62)
      \lstinputlisting{programs/fun-not-val.txt}                                 & {
          \begin{tabularx}{1in}{r l}
            83\%   & \verb|4|     \\
            **11\% & \verb|error| \\
            4\%    & \verb|2|     \\
            1\%    & \verb|8|     \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{IsolatedFun} Functions can't refer to free variables
      except for the built-in ones.
                                                                                 &
      % scope.refer_global_is_possible (5)
      \lstinputlisting{programs/isolated-fun.txt}                                & {
          \begin{tabularx}{1in}{r l}
            77\%   & \verb|3|     \\
            **23\% & \verb|error| \\
          \end{tabularx}
        }
      \\\hline
      \miscon{NoCircularity} Data structures can't (possibly indirectly)
      refer to themselves.                                                       &
      % vector2.warmup_circularity (53)
      \lstinputlisting{programs/no-circularity.txt}                              & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            76\%   & \verb|3|                   \\
            **14\% & \verb|error|               \\
            9\%    & Run out of memory or time. \\
            1\%    & \verb|+inf|                \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{StructByRef} Data structures might refer to variables by their references.                                                       &
      % vector2.alias_var_in_mvec (49)
      \lstinputlisting{programs/struct-by-ref.txt}                              & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            67\%   & \verb|#(1 2 3)|                   \\
            **24\% & \verb|#(1 2 4)|               \\
            9\%    & error \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{StructsCopyStructs} Storing data structures into data structures makes copies.                                                       &
      % lambda1.smol_quiz_circularity (58)
      \lstinputlisting{programs/structs-copy-struct.txt}                              & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            65\%   & \verb|#0=#(#0# #0#)| (Racket circular object notation) \\
            24\%   & \verb|#(#(2 3) #(2 3))|               \\
            **6\%   & \verb|#(#(2 #(2 3)) #(2 3))|                   \\
            6\%    & error \\
          \end{tabularx}
      }                                                                                   \\\hline
    \end{tabularx}
    \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
      with ``**'' represent the misconception. (Part II)}
    \label{t:tutor-grounded-misconceptions-2}
  \end{table}

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
    \textbf{Misconception}                                                             &
    \textbf{Question}                                                                  &
    \textbf{Table of Answers}                                                              \\\hline
    \surmisedMiscon{NestedDef} Sequences of definitions are
    interpreted as if they are written in nested blocks. A definition
    is not in the scope of later definitions.
    &
    % program: scope.what_is_x_4 (13)
    \lstinputlisting{programs/nested-def.txt}                                          & {
        \begin{tabularx}{1in}{r l}
          92\%  & \verb|2| \\
          **8\% & \verb|1| \\
        \end{tabularx}
    }                                                                                      \\\hline
    \surmisedMiscon{Lazy} Expressions are only evaluated when their values are needed. &
    % order.bad_order (20)
    \lstinputlisting{programs/lazy.txt}                                                & {
        \begin{tabularx}{1in}{r l}
          **57\% & \verb|1 3|   \\
          43\%   & \verb|error| \\
        \end{tabularx}
    }                                                                                      \\\hline
  \end{tabularx}
  \caption{Surmised misconceptions identified by the SMoL~Tutor. Answers marked
    with ``**'' represent the misconception.}
  \label{t:tutor-surmised-misconceptions}
\end{table}

\paragraph{Confirmed Misconceptions}

The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.

\paragraph{Added Misconceptions}

The misinterpreters helped us find misinterpretations that we had
overlooked. For instance, consider this program from \quizState:
\begin{lstlisting}
(defvar x 12)
(deffun (f x)
  (set! x 0))
(f x)
x
\end{lstlisting}
We had assumed the wrong answer
\texttt{0} is caused by \miscon{CallByRef}.
However, our misinterpreters made us realize it can also be explained
by \miscon{FlatEnv}. This could also explain why we see a difference in
error rate when the formal and actual parameter have the same
name (as mentioned in \Cref{s:misconceptions-quiz:state}).

% Tutor's refined program sets and misinterpreters also helped us to
% identify programs that detect one of the misconceptions but not the
% other. Consider this new program
% % mut-vars.not_aliased_by_defvar_1 (29)
% \begin{lstlisting}
%   (defvar x 12)
%   (defvar y x)
%   (set! x 0)
%   x
%   y
% \end{lstlisting}
% 12\% of students asserted that the following program produces
% \texttt{0 0}, which suggest that they have the \textbf{BindByRef}
% misconception. Students with a \miscon{FlatEnv} misconception would
% have chosen \texttt{0 12}, which happens to be the correct answer.
% (1\% wrote \textit{depends on
% implementation}; 2\% asserted \texttt{error}; the remaining 85\%
% asserted \texttt{0 12}).

% The second program in \Cref{t:quizzes-data-fun} is already a good
% detector for \miscon{FlatEnv}. SMoL Tutor includes an alternative
% program that does not depend on equality checks.
% % scope.error_when_refer_to_undefined (9)
% \begin{lstlisting}
%   (deffun (addy x)
%     (defvar y 200)
%     (+ x y))
%   (+ (addy 2) y)
% \end{lstlisting}
% 4\% of students asserted that the following program produces
% \texttt{402}, which suggest that they have the \miscon{FlatEnv}
% misconception. Students with a \textbf{BindByRef} misconception would
% have chosen \texttt{error}, which happens to be the correct answer.
% (the remaining 96\% asserted \texttt{error}).

\paragraph{A New Potential Misconception}

For the following program, added in the Tutor:
\begin{lstlisting}
  (defvar y (+ x 2))
  (defvar x 1)
  x
  y
\end{lstlisting}
56\% of students asserted that it produces
\texttt{1 3}. Based on this, we surmise that students might have
another misconception, which we define as \surmisedMiscon{Lazy}. (Recall that
SMoL is eager, but even in many lazy languages, this would be an error.)

\paragraph{Another Potential Misconception, and Its Effect on
  Interpreting Descriptions.}%
\label{ss:misinterpreters:submisconception}

Consider this program from the SMoL Tutor:
\begin{lstlisting}
(defvar x 1)
(deffun (main)
  (deffun (get-x) x)
  (defvar x 2)
  (get-x))

(main)
\end{lstlisting}
This program, suitably translated, would produce \verb|2| in a wide
variety of languages (Python, JavaScript, Racket, Java, etc.), because
\verb|get-x| and the inner \verb|x| are in the same scope block. The
answer \verb|1| cannot be explained by any of our existing
misconceptions. Based on this, we surmise a new misconception,
\surmisedMiscon{NestedDef}. (Though its frequency falls below our
threshold, our reading of answers suggests this may be more
widespread, and we feel it needs to be investigated more.)

Once we turned this into a misinterpreter, we found that it
unexpectedly captured the program-answer pair for the following
program from \quizFun---which should be an error, due to the
double-binding of \verb|x| in the same scope block---and the answer
\verb|2 0|:
\begin{lstlisting}
  (defvar x 0)
  (defvar y x)
  (defvar x 2)
  x
  y
\end{lstlisting}
Previously, we had interpreted this only as \miscon{DefOrSet}, because
students had stated that the second \lstinline|(defvar x ....)|
``mutates'' or ``redefines'' \verb|x|. This is reminiscent of the
behavior of languages like Python, which use the same syntax both for
binding new variables and for mutating existing ones.

The problem here is that the word ``redefine'' underspecifies how the
second definition is interpreted. We had interpreted it as
\textit{mutating} the binding established by the first definition,
which fits \miscon{DefOrSet}. However, another possibility is that
it \emph{shadows} the binding (i.e., establishes a new scope block). We
did not recognize that the original misconception is underspecified
until we uncovered the new (surmised) misconception.

\paragraph{Summary}

To summarize, we used the SMoL Tutor, with an expanded set of
programs, to investigate student misconceptions. We implemented the
idea of misinterpreters to help us properly classify student
performance. We were able to reconfirm all the previously identified
misconceptions, refine some of them, and also identify potential new
ones (that need further investigation).

\section{Is The Tutor Effective?}
\label{s:effect-tutor}

\begin{figure}[t]

  \includegraphics[width=1\textwidth]{images/Misconceptions_Decay.png}

  \caption{How many students chose a wrong answer that (uniquely)
    represents a misconception? (Downward tendency suggests improvement
    over time.)}
  \label{f:tutor-trend-plots}

\end{figure}

Recall that the SMoL Tutor is not only a collection of MCQs: it is
also a \emph{tutor}! So far, we have investigated the value of the
MCQs. Now we examine its tutoring aspect. Concretely, we ask:
\begin{description}

\item[RQ] How effective is the Tutor at correcting each misconception?

\end{description}
To study this, we perform the following analysis per
misconception. Using misinterpreters, we identify those questions
whose wrong answers fit only one misconception (i.e., only one output
matches that produced by that misinterpreter, and it matches only that
misinterpreter). We then examine student performance over time across
those questions. This would let us examine how they do on just that
topic, in isolation, over time.

We present the result of this analysis in two forms. Graphically, we
show plots in \Cref{f:tutor-trend-plots}. Each figure shows the percentage of students who
chose the answer corresponding to that misconception. Ideally, we
would like to see these percentages diminish.

Indeed, that is what we see in most of the graphs. The exceptions are
\miscon{CallsCopyStructs}, \surmisedMiscon{NestedDef}, and
\miscon{StructsCopyStructs}, which have only one problem (and hence no
trend), and \miscon{DefOrSet} and \miscon{FunNotVal}, which show an
increase. The lack of improvement for \miscon{FunNotVal} is
unsurprising because the Tutor does not explicitly address this issue,
focusing on closures created by \lstinline|lambda|s rather than named
functions. (However, this does not explain the increase!)

We also perform a logistic regression to see whether these
improvements are significant (at a $p < 0.05$ threshold); details are
in \supplRef{Paper.html}. Of the 11:
\begin{itemize}

\item Of the nine seemingly improved (i.e., decreased)
  misconceptions:
  \begin{itemize}
  \item Two are \emph{not} significant: \miscon{CallByRef}
      (\pValue\ = .074); \miscon{NoCircularity} (\pValue\ = .075).
    \item The other seven \emph{are} significant.
    \end{itemize}
  \item However, the two with an increasing trend (\miscon{DefOrSet}
    and \miscon{FunNotVal}) are \emph{also} significant.
\end{itemize}

These data broadly suggest that the Tutor is a net positive.
Ultimately, however, what these data really show is a need for
improvement in the Tutor. When designing the Tutor questions, many
more were intended to be representative of single
misconceptions. However, as noted in \Cref{s:misinterp-intro}, it is
easy to be incomplete (or incorrect) in ascribing
misconceptions. Furthermore, as their set grows, it is difficult to
reassess all the problems manually. Once evaluated using
misinterpreters, we found far fewer problems than we would have liked.
Concretely, only 40 of the 71 eligible problems (after removing the
non-SMoL modules) were useful in the above
analysis.

We therefore view the above data as
purely formative: they suggest that the Tutor \emph{most likely did
  not do harm} and perhaps even \emph{may have done some
  good}. However, it would be improper to read too much into the
analysis. It is quite possible that some of the other problems would
have found issues. Rather, what we really see is value in the
misinterpreter concept. It is not only useful for analysis, it is also
valuable for \emph{problem design}: in the next iteration of the
Tutor, we will use misinterpreters actively to shape the incorrect answers and,
as necessary, update the programs as well. We therefore hope to have
much more thorough analyses in future work.

\section{Performance on Other Populations}
\label{s:generality-tutor}

There is, of course, a significant danger that the data above have
been overfitted to only one institution (\Cref{s:univ-1-course},
henceforth University 1),
thereby actually reflecting the state of its curriculum rather than
some greater truth about program behavior understanding.  We already
have some reason to believe this is not the case: the related work
discussed in \Cref{s:rel-work} is drawn from many institutions in
multiple countries with different educational preparations, levels,
and demographics. Nevertheless, that gives us only
limited information about the specific questions and misconceptions
described above.

Fortunately, we were able to deploy the Tutor on two other
populations:
\begin{itemize}

\item University 2 is a primarily public university in the US. It is
  one of the largest Hispanic-serving institutions in the country. As
  such, its demographic is extremely different from those whose data
  were used above. The Tutor was used in one course in Spring 2023,
  taken by 12 students.
  The course is a third-year, programming
  language course. The students are required to have taken two
  introductory programming courses (C++ focused).

\item A separate instance of the Tutor was published on the
  website of a programming languages textbook [ANON]. Over the course
  of 8 months, 597 people started with the first module and 103
  users made it to the last one. To protect privacy, we intentionally
  do not record demographic information, but we conjecture that the
  population is largely self-learners  (who are known to use the
  accompanying book),
  including some professional
  programmers. It is extremely
  unlikely to be the students from either university, because they
  would not get credit for their work on the public instance; they
  needed to use the university-specific instance. Furthermore, since
  they were not penalized for wrong answers, it would make little
  sense to do a ``test run'' on the public instance. Finally, we note
  that there is no overlap between the dates of submission on the
  public instance and the semester at University 1.

\end{itemize}
These two populations are therefore at least somewhat different from
the original population, and help us assess whether the problems we
identify are merely an artifact of the first institution.

To evaluate, we computed a Spearman's rank correlation $\rho$, ranking
questions by what percentage of students got the question
right. Between the original university and University 2,
we obtain a \pValue~=~2.013e-07. Between the original university and
the online population, we obtain a \pValue~<~2.2e-16. These show that
the other two populations performed similarly to the original one.
While
further validation on other populations remains essential, these
suggest that the questions are \emph{finding misconceptions that may
  be universal}.

\section{Related Work}
\label{s:rel-work}

\begin{table}[t]
  \centering
  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.5\hsize\RaggedRight}X >{\hsize=1.4\hsize\RaggedRight}X}
    \textbf{Publication}                       &
    \textbf{Population}                       &
    \textbf{Languages}                       &
    \textbf{Misconceptions}
    \\\hline
    \cite{fleuryParameterPassingRules1991} &
    Unclear, likely CS2 students &
    Pascal &
    \surmisedMiscon{CallerEnv}; \miscon{IsolatedFun}; \miscon{DefOrSet}
    (See their RULEs 2, 3a, and 3b in TABLE 2;
    RULE 1 doesn't apply to us.)
    \\\hline
    \cite{goldmanIdentifyingImportantDifficult2008,goldmanSettingScopeConcept2010} &
    CS1 students &
    Java, Python, Scheme &
    Scope and memory model (See their Figures 4 and 5)
    \\\hline
    \cite{fislerAssessingTeachingScope2017} &
    Third- and fourth-year undergrads &
    Java and Scheme &
    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{CallsCopyStruct}; \miscon{DefByRef}
    (See their Section 4)
    \\\hline
    \cite{saarinenHarnessingWisdomClasses2019} &
    CS2 students &
    Java &
    \miscon{StructByRef} (Their G2); \miscon{DefByRef} or \miscon{DefCopyStructs} (Their G3); \miscon{CallsCopyStruct} (Their G4).
    \\\hline
    \cite{strombackProgressionStudentsAbility2023} &
    CS masters &
    Python &
    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
    (See their section 4.2)
    \\\hline
    \cite{strombackProgressionStudentsAbility2023} &
    CS undergrads &
    C++ &
    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
    (See their section 4.2)
    \\\hline
  \end{tabularx}
  \caption{Similar misconceptions found in prior research.}
  \label{t:rel-work}
\end{table}

\paragraph{Misconceptions}

Misconceptions related to scope, mutation, and higher-order functions have been
widely identified in varying populations
(from CS1 students to graduate students to users of online forums)
over many years (since at last 1991)
in varying programming languages (from Java to Racket) and in
different countries (such as the USA and Sweden).
\Cref{t:rel-work} lists works that seem most relevant to us.

There are some small differences.
\cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
that is different from \miscon{FlatEnv}:
\begin{description}
  \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
  When a function is called, the function body is evaluated in an
  environment that extends from the \textit{caller's} environment.
\end{description}
We don't include \surmisedMiscon{CallerEnv} in our analysis because in our
data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
are also explainable by \miscon{FlatEnv}.

Appendix A of \cite{sorvaVisualProgramSimulation2012} provides an
extensive survey of misconceptions reported
in research up to 2012.
There are overlaps between our survey and
theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
However, because their descriptions are brief, it is difficult to tell
whether a misconception in their survey matches our misconceptions.
Because
they provide neither misinterpreters nor representative program-output
pairs, it is difficult to determine the overlaps precisely (showing
the value of providing these two machine-runnable descriptions).
At any rate,
we certainly find
no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
\miscon{DefByRef} in their survey.

\paragraph{Goal Sentences}

Our idea of ``goal sentences'' is not substantially different from the
\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
We only used these sentences (or rules) to guide the
design of the Tutor and as text in the Tutor itself.
\cite{duranRulesProgramBehavior2021} argue for other uses of such
sentences.

\paragraph{Misinterpreters}

Our idea of misinterpreters is related to mystery languages
\cite{diwanPLdetectiveSystemTeaching2004,
pombrioTeachingProgrammingLanguages2017}. Both approaches use
evaluators that represent alternative semantics to the same syntax.
However, the two are complementary.
In mystery languages, instructors design the space of semantics with
pedagogic intent, and
students must create programs to explore that space. Misinterpreters,
in contrast, are driven by student input, while the programs are provided
by instructors. The two approaches also have different goals:
mystery languages focus on encouraging students to experiment with
languages; misinterpreters aim at capturing students' misconceptions.

\section{Threats to Validity}
\label{s:threats}

\subsection{Construct Validity}

Did we measure the right thing?
While our goal is to study student understanding of language
behavior, what we actually measure is performance on MCQs on select
programs. MCQs as a mechanism introduce various clear biases, though
we add the ``Other'' options to somewhat alleviate them. The small
size, syntactic details (such as intentionally meaningless variable
names), and other aspects of the programs may also impact our ability
to measure understanding. (This could go both ways: students may have
no trouble understanding behavior in a small program but may struggle
to do so in a larger one.)

A particularly notable threat is the use of Lispy
syntax. We chose it for two reasons: both because it helps clarify
scope (\Cref{s:smol}) and because the rest of the course used
Racket. However, students may well perform differently with more
familiar syntaxes. It seems unlikely their performance would be \emph{too}
different given the many languages that have produced similar
misconceptions (\Cref{t:rel-work}). Nevertheless, we intend to use a
variety of syntaxes to examine this issue further.

Finally, we curated programs by hand, so they may well reflect our own
biases about misconceptions. It may be possible to mitigate this
problem by synthesizing MCQ programs using the misinterpreters.

\subsection{Internal Validity}

Is our reasoning valid? We have applied standard techniques and
measurements for evaluating student responses. However, our
instruments still lack the validity of a proper concept
inventory. Their creation requires heavyweight processes (such as
Delphi methods \cite{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
well as conversations with learners, and can hence be prohibitive in
cost. The Quizius method (\Cref{s:quizius}) was created precisely to
be an inexpensive method that provides a good proxy.

Ultimately, our goal is to provide a reasonable instrument for
widespread use. While the set of all misconceptions could be
unbounded, we believe our tasks, especially as embedded in the Tutor,
provide a good starting point for others. In particular, if students
select a wrong answer, that is still of \emph{some} use to an
educator, even if the precise misconception cannot be pinned down with
the highest accuracy. We therefore believe our instruments, and our
misinterpreter technique, are of general value.

A failure in our Tutor's logging infrastructure led us to miss some
responses. These are unlikely to be task-specific because the Tutor
has a generic framework that should perform the same across
tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, we do not believe
this had a noticeable impact. (In addition, every wrong answer is
still wrong! We may just not have exactly the right proportions of
them.)

\subsection{External Validity}

How well do our results generalize? Certainly there is reason to
question whether our results would apply to other
populations. \Cref{s:generality-tutor} provides preliminary evidence
that the results are not specific to one institution, and
\Cref{s:rel-work} suggests these issues are widespread. Nevertheless,
much more broad testing is needed to confirm our specific instruments.

The other major concern is the tie to Lispy syntax. Learners in other
settings may do worse or even better with it. Building a Tutor that
supports multiple, and more traditional, syntaxes should help address
this issue. We defer this to future work.

\section{Discussion}%
\label{s:discussion}

The paper has already identified several areas for future work:
\begin{itemize}
\item testing on more varied populations;
\item using other syntaxes;
\item having more questions that uniquely identify minsconceptions;
  and,
\item enabling textual responses in the Tutor so we can better
  characterize wrong answers.
\end{itemize}
These can help get us even closer to a good approximation of a true concept
inventory. We would also like the Tutor to make more use of the education
theories discussed in \Cref{s:background}.

We focus here on some issues that we think warrant broader
discussion.

\paragraph{The State-Aliasing-Function Triangle}

Our problematic programs hardly include any ``advanced'' programming
features: there are no threads, asynchrony, sophisticated type systems, ownership,
inversion of control, etc. Indeed, many of those features typically
build on an understanding of these basics (e.g., it is hard to make
sense of ownership \cite{clarkeOwnershipTypesFlexible1998} without a good understanding of this
triangle). But many populations seem to struggle even with this, which
may explain why concepts like ownership are considered hard \cite{crichtonGroundedConceptualModel2023}.

It is worth noting that we find problems even without all three
components, as \Cref{s:smol-quizzes} shows! Nevertheless, we think it
would be useful for curricula to focus on achieving mastery of this
triangle. This may require a deep revision of widely accepted
pedagogy. For instance, it is common to explain variables as
``boxes''. But if taken seriously by a learner, this metaphor may
cause more harm than good. As prior research \cite{putnamSummaryMisconceptionsHigh1986,groverMeasuringStudentLearning2017,hermansThinkingOutBox2018} shows, students expect that a variable can then contain more
than one value, removing one makes the other accessible, etc.  That
research has not explored aliasing, but the metaphor may affect that
too. A box is a closed object, so a (mutable) value put ``into'' it
clearly can't be modified by another ``box'' (variable)---i.e., it not
only doesn't explain but is antithetical to aliasing.

\paragraph{Terminology}

It is common to teach programming using the terms ``call-by-value''
and ``call-by-reference''. The reader will note that we instead have
\emph{three} \miscon{ByRef} misconceptions.

The emphasis on ``calling'' suggests that the semantics is associated
\emph{only} with calls. That leaves open what happens when one simply
binds a variable. In a reasonable language (and definitionally, in
SMoL), the behavior is exactly the same: the formal parameter can be
viewed as a binding to the value of the actual. But students often
form inconsistent views because the ``call'' terminology breaks this
deep similarity. We therefore recommend that languages in general use
the term \emph{bind-by-}, to emphasize the underlying semantic unity
of these syntactically different mechanisms.

We feel that further confusion is caused by terms like ``pass'' and
``return''. We often vocalize the call \verb|(f x)| as ``passing
\verb|x| to \verb|f|''; saying ``passing \emph{the value of} \verb|x|
to \verb|f|'' is a mouthful. But is it then surprising that students
think \verb|x| is being aliased? Similarly, consider a statement like
\verb|return y| (in Python syntax). Of course, semanticists understand
that it is the \emph{value} of \verb|y|, not \verb|y| itself, that is
being returned. Nevertheless, it is not surprising if this also
results in assumptions that \verb|y| is either aliased or that it has
escaped from the function (leading to an interpretation of dynamic
scope). We believe there is a need for significant research to
investigate these kinds of effects, which are similar to but not
strictly the same as ``vernacular misconceptions'' \cite{nationalresearchcouncilScienceTeachingReconsidered1997}.

\end{document}